---
title: "p8105_hw2_xc2473"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
```
##Quesiton 1
```{r, message = FALSE}
library(tidyverse)

library(readxl)

mtw_data = 
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, col_names = TRUE) %>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered) %>% 
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(sports_balls)) 
  
```
```{r, message = FALSE}
precip_17_data =
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, col_names = TRUE, skip = 1) %>% 
  janitor::clean_names() %>%
  drop_na(total, month) %>%
  mutate(year = 2017) %>% 
  select(year, month, total)

precip_18_data =
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, col_names = TRUE, skip = 1) %>% 
  janitor::clean_names() %>%
  drop_na(total, month) %>%
  mutate(year = 2018) %>% 
  select(year, month, total)

precip_1718_data = 
  full_join(precip_17_data, precip_18_data, by = c("year", "month", "total")) %>% 
  group_by(year) %>% 
  mutate(month = month.name[month],
         year_total = sum(total) )
```
There are `r nrow(mtw_data)` observations in mtw_data dataset. Dumpster, weight_tons and volume_cubic_yards are examples of key variables. There are `r nrow(precip_1718_data)` obserrvations in precip_1718_data dataset.Year, month and total are key variables of it.
The total precipitation in 2018 was `r sum(pull(precip_18_data, total))`. The median number of sports balls in a dumpster in 2017 was `r median(pull(mtw_data %>% filter(year==2017), sports_balls))`.

###Question 2###
```{r, message = FALSE}
pols_month_data = read_csv(file = "./fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "([\\-])") %>%
  mutate(year = as.numeric(year),
         month = month.abb[as.numeric(month)],
         day = as.numeric(day),
         president = ifelse(prez_dem == 1, "DEM", "GOP")
         ) %>% 
  select(-prez_dem, -prez_gop, -day) %>%
  arrange(year, month)
  
snp_data = read_csv(file = "./fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), sep = "([\\/])") %>%
  mutate(year = as.numeric(year),
         month = month.abb[as.numeric(month)],
         day = as.numeric(day),
         ) %>% 
  select(year, month, day, close) %>%
  arrange(year, month)

unempolyment_data = read_csv(file = "./fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>% 
  rename(Jan = jan, Feb = feb, Mar = mar, Apr = apr, May = may, Jun = jun, Jul = jul, Aug = aug, Sep = sep, Oct = oct, Nov = nov, Dec = dec) %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment") 

merge_1_data = left_join(pols_month_data, snp_data, by = c("year", "month"))

merge_data = left_join(merge_1_data, unempolyment_data, by = c("year", "month"))
```

There are `r nrow(pols_month_data)` observations and `r ncol(pols_month_data)` varaibles in pols_month_data dataset. The range of years is `r range(pull(pols_month_data, year))`. The key variables are year, month, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem, and presidents.
There are `r nrow(snp_data)` observations and `r ncol(snp_data)` varaibles in snp_data dataset. The range of years is (`r range(pull(snp_data, year))`). The key variables are year, month, day and close.
There are `r nrow(unempolyment_data)` observations and `r ncol(unempolyment_data)` varaibles in pols_month_data dataset. The range of years is (`r range(pull(unempolyment_data, year))`). The key variables are year, month, unemployment.
There are `r nrow(merge_data)` observations and `r ncol(merge_data)` varaibles in the resulting dataset. The range of years is (`r range(pull(merge_data, year))`). The key variables are year, month, unemployment, close, gov_gop, sen_gop, rep_gop, gov_dem and so on.

###Question 3###
Load and tidy the data.
```{r, message = FALSE}
pop_baby_name_data = read_csv(file = "./Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  mutate(childs_first_name = str_to_upper(childs_first_name),
         ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC")) %>% 
  distinct()
```
Rank in popularity of the name “Olivia” as a female baby name over time
```{r, message = FALSE}
olivia_rank = pop_baby_name_data %>% 
  filter(childs_first_name == "OLIVIA", gender == "FEMALE") %>% 
  select(childs_first_name, gender, ethnicity, year_of_birth, rank) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  ) 
knitr::kable(olivia_rank)
```

A table showing the most popular name among male children over time.

```{r, message = FALSE}
most_pop_male_name = pop_baby_name_data %>% 
  filter(gender == "MALE", rank == 1) %>% 
  select(childs_first_name, ethnicity, year_of_birth) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name
  )  
  knitr::kable(most_pop_male_name)
```

For male, white non-hispanic children born in 2016, a scatter plot shows the number of children with a name (y axis) against the rank in popularity of that name (x axis)

```{r, message = FALSE}
wnh_2016 = pop_baby_name_data %>% 
  filter(gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016 )

ggplot_wnh_2016 = ggplot(wnh_2016, aes(x = rank, y = count)) + geom_point()

ggplot_wnh_2016
```

